---
title: "CACTES COVID Analysis"
output:
  pdf_document: default
  word_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)   # contains kable() function
#setwd("~/Documents/CACTES")
#loading data
data <- read_csv("Combined Cleaned Data.csv")
```


**Descriptive statistics – ethnicity**
```{r}
#switch letter option with the name of ethnicity
data$Q1[data$Q1 == "A"] <- "Bunyankole"
data$Q1[data$Q1 == "B"] <- "Muganda"
data$Q1[data$Q1 == "C"] <- "Mukiga"
data$Q1[data$Q1 == "D"] <- "Musoga"
data$Q1[data$Q1 == "E"] <- "Muhaya"
data$Q1[data$Q1 == "F"] <- "Other"

#code to print bar graph
ggplot(data, aes(x = Q1)) + geom_bar(fill="steelblue") + labs(x = "Ethnicity", y = "Count")

```


**Descriptive statistics – What are the 3 most common ethnicities? ** 
```{r}
#count unique data points for Question 1
Q1_table <- data %>%
  count(Q1, sort = TRUE)
total <- dim(data)[1] #total number of observations
Q1_table$Percentage <- (Q1_table$n/total)*100
Q1_table <- Q1_table %>%
  rename(Ethnicity = Q1,
         Count = n)
print(Q1_table)
```
Top three ethnicities are Muganda(85.27%), Other (10.08%), and Bunyankole (2.325%)


**Descriptive statistics – religion**
```{r}
#switch letter option with the name of religion
data$Q2[data$Q2 == "A"] <- "Christian"
data$Q2[data$Q2 == "B"] <- "Christian"
data$Q2[data$Q2 == "C"] <- "Christian"
data$Q2[data$Q2 == "A,B"] <- "Christian"
data$Q2[data$Q2 == "B,C"] <- "Christian"
data$Q2[data$Q2 == "D"] <- "Muslim"
data$Q2[data$Q2 == "E"] <- "Born Again"
data$Q2[data$Q2 == "F"] <- "No religion"
data$Q2[data$Q2 == "G"] <- "Other"

#code to print bar graph
ggplot(data, aes(x = Q2)) + geom_bar(fill="steelblue") + labs(x = "Religion", y = "Count")


#code to print bar graph
ggplot(data, aes(x = Q2)) + geom_bar(fill="steelblue") + labs(x = "Religion", y = "Count")
```

**Descriptive statistics – What is the gender breakdown? **

```{r}

data$Female[data$Female == 1] <- "Female"
data$Female[data$Female == 0] <- "Male"

#code to print bar graph
ggplot(data, aes(x = Female)) + geom_bar(fill="steelblue") + labs(x = "Religion", y = "Count")

#rename the table with appropiate column names
Female <- data %>%
  count(Female, sort = TRUE) %>%
  rename(Count = n) %>%
  mutate(Percentage = (Count/sum(Count))*100)

print(Female)

```


**Descriptive statistics – What are the 3 most common religions? **
```{r}
Q2_table <- data %>%
  count(Q2, sort = TRUE)

Q2_table$Percentage <- (Q2_table$n/total)*100
Q2_table <- Q2_table %>%
  rename(Religion = Q2,
         Count = n)
print(Q2_table)
```
Three most common religions are Catholic (45.74%), Muslim (37.2%), and Anglican (6.2%)


**Descriptive statistics – Education**
```{r}
data$Q3[data$Q3 == "A"] <- "Primary"
data$Q3[data$Q3 == "B"] <- "Secondary"
data$Q3[data$Q3 == "C"] <- "Tertiary"
data$Q3[data$Q3 == "D"] <- "Bachelors and Higher"
data$Q3[data$Q3 == "E"] <- "Bachelors and Higher"
data$Q3[data$Q3 == "F"] <- "None"

#factor the options so they go in desired order
data$Q3 <- factor(data$Q3, levels = c("None", "Primary", "Secondary", "Tertiary", "Bachelors and Higher"))


#code to print bar graph
ggplot(data, aes(x = Q3)) + geom_bar(fill="steelblue") + labs(x = "Education", y = "Count")
```


*Descriptive statistics – what is the percent of respondents that held the highest level of education? **

```{r}
Q4_table <- data %>%
  count(Q4)

Q4_table$Percentage <- (Q4_table$n/total)*100
Q4_table <- Q4_table %>%
  rename(Highest.Level = Q4,
         Count = n)
print(Q4_table)
```
30.23% of respondents have the highest level of education in his/her household. 68.99% are not.


**Descriptive statistics – What is the [average/media etc.] household size?**

```{r}
Q6_vector <- data$Q6

#calculate mean
Q6_mean <- mean(Q6_vector, na.rm = TRUE) #remove anyone who didn't answer this question
print(Q6_mean)

#calculate median
Q6_med <- median(Q6_vector, na.rm = TRUE) #remove anyone who didn't answer this question
print(Q6_med)


#create function for mode
mode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
Q6_mode <- mode(Q6_vector)
print(Q6_mode)

```


**What is the average percent of household members that generate an income?**
```{r}
names <- c("Q6", "Q10")
Household_data <- data[names]
#create percent of workers in each household
Household_data <- Household_data %>%
  mutate(Percent.of.Workers = (Q10/Q6)*100)
#average percent 
worker_mean <- mean(Household_data$Percent.of.Workers, na.rm = TRUE) #remove any observations that were NA

print(worker_mean)
```

The average percent of household members that generate an income is `worker_mean` 


**Descriptive statistics – number of children in the household under the age of 18**
```{r}
Q7_vector <- data$Q7
#calculate mean
Q7_mean <- mean(Q7_vector, na.rm = TRUE) #remove anyone who didn't answer this question
#calculate median
Q7_med <- median(Q7_vector, na.rm = TRUE) #remove anyone who didn't answer this question
#calculate mode
Q7_mode <- mode(Q7_vector)
```
Mean = `Q7_mean`
Median = `Q7_med`
Mode = `Q7_mode`


**Descriptive statistics – What is the percent of children (<18) in the household that attends school regularly before COVID19**
```{r}
names.2 <- c("Q7", "Q8")
Children_data <- data[names.2]

#create percent of workers in each household
Children_data <- Children_data %>%
  mutate(Percent.of.Children = (Q8/Q7)*100)

#average percent 
children_mean <- mean(Children_data$Percent.of.Children, na.rm = TRUE) #remove any observations that were NA

print(children_mean)
```




**Descriptive statistics – What is the (average/median) number of people working in a household? **
```{r}
Q10_vector <- data$Q10
#calculate mean
Q10_mean <- mean(Q10_vector, na.rm = TRUE) #remove anyone who didn't answer this question
#calculate median
Q10_med <- median(Q10_vector, na.rm = TRUE) #remove anyone who didn't answer this question
#calculate mode
Q10_mode <- mode(Q10_vector)
```
Mean = `Q10_mean`
Median = `Q10_med`
Mode = `Q10_mode`


**Descriptive statistics – What sectors does the sample population work in?**
```{r}
sectors <- c("Agriculture", "Transportation", "Education" ,"Sponsership", "Construction", "Business", "Trade", "Government", "Law", "Health", "Food", "Other")

count <- c(sum(data$agriculture),sum(data$transportation), sum(data$education), sum(data$sponsership), sum(data$construction), sum(data$business), sum(data$trade), sum(data$government), sum(data$law), sum(data$health), sum(data$food), sum(data$other))

Work_data <- tbl_df(cbind(sectors, count))
Work_data$count <- as.numeric(Work_data$count)

Work_data <- Work_data %>%
  mutate(Percent.Work = (count/total)*100)
print(Work_data)
```


**Are there any differences in sectors that people work in between the two districts? **
```{r}
#graph of 2 districts 
Kalungu <- data %>%
  filter(District == "KALUNGU") %>%
  select(District, agriculture, transportation, education, sponsership, construction, business, trade, government, law, health, food, other)
Masaka <- data %>%
  filter(District == "MASAKA") %>%
  select(District, agriculture, transportation, education, sponsership, construction, business, trade, government, law, health, food, other)

#melt data to more friendly format
Kalungu_melt <- melt(Kalungu, measure.vars = c("agriculture", "transportation", "education", "sponsership", "construction", "business", "trade", "government", "law", "health", "food", "other"))
Masaka_melt <- melt(Masaka, measure.vars = c("agriculture", "transportation", "education", "sponsership", "construction", "business", "trade", "government", "law", "health", "food", "other"))


Kalungu_melt <-  Kalungu_melt %>%
  group_by(variable) %>%
  summarise(Total = sum(value)) %>%
  rename(Sector = variable, Count = Total)

Masaka_melt <-  Masaka_melt %>%
  group_by(variable) %>%
  summarise(Total = sum(value))%>%
  rename(Sector = variable, Count = Total)

ggplot(data=Kalungu_melt, aes(x=Sector, y=Count)) + geom_bar(stat="identity", fill = "royalblue4") + labs(x = "Sector", y = "Total") + theme_bw() +
  theme(axis.text.x=element_text(size=rel(1.1), angle=90))+ ggtitle("Sector Overview in Kalungu Region")

ggplot(data=Masaka_melt, aes(x=Sector, y=Count)) + geom_bar(stat="identity", fill = "royalblue4") + labs(x = "Sector", y = "Total") + theme_bw() +
  theme(axis.text.x=element_text(size=rel(1.1), angle=90))+ ggtitle("Sector Overview in Masaka Region")

#Kalungu region
print(Kalungu_melt)

#Masaka region
print(Masaka_melt)
```

**What is the average monthly income?**
```{r}
Q12_vector <- data$Q12
#calculate mean
Q12_mean <- mean(Q12_vector, na.rm = TRUE) #remove anyone who didn't answer this question
print(Q12_mean)

#calculate standard deviation
Q12_sd <- sd(Q12_vector, na.rm = TRUE)#remove anyone who didn't answer this question
print(Q12_sd)
```
Mean = `Q12_mean`
Standard Deviation = `Q12_sd`


**What is the daily expenditure before lockdown? **
```{r}
Q13_vector <- data$Q13
#calculate mean
Q13_mean <- mean(Q13_vector, na.rm = TRUE) #remove anyone who didn't answer this question
print(Q13_mean)

#calculate standard deviation
Q13_sd <- sd(Q13_vector, na.rm = TRUE)#remove anyone who didn't answer this question
print(Q13_sd)
```
Mean = `Q13_mean`
Standard Deviation = `Q13_sd`


**What is the daily expenditure after lockdown?**
```{r}
Q14_vector <- data$Q14
#calculate mean
Q14_mean <- mean(Q14_vector, na.rm = TRUE) #remove anyone who didn't answer this question
print(Q14_mean)

#calculate standard deviation
Q14_sd <- sd(Q14_vector, na.rm = TRUE)#remove anyone who didn't answer this question
print(Q14_sd)
```
Mean = `Q14_mean`
Standard Deviation = `Q14_sd`


**How does the level of education of the respondent affect the economic status of a household? **
```{r}
education.lm <- lm(Q13 ~ Q3, data = data)
summary(education.lm)
```


**How does of the level of education of the respondent affect the change in economic status of the household post-COVID19**
```{r}
#add new column 
Difference <- data %>% 
  select(Q3, Q13, Q14) %>%
  mutate(Change = Q14 - Q13)

difference.lm <- lm(Change ~ Q3, data = Difference)
kable(coef(summary(difference.lm)), digits = c(3,3,3,3))
```


**What are the top 3 household expenses? **
```{r}
Groups <- c("Food", "PPE", "Medicine" ,"Water", "Electricity", "Fuel/charcoal", "Education", "Healthcare", "Other")

Count <- c(sum(data$Q15.A),sum(data$Q15.B),sum(data$Q15.C),sum(data$Q15.D),sum(data$Q15.E),sum(data$Q15.F),sum(data$Q15.G),sum(data$Q15.H), length(which(!is.na(data$Q15_Other))))

Expense_data <- tbl_df(cbind(Groups, Count))
Expense_data$Count <- as.numeric(Expense_data$Count)
Expense_data <- Expense_data[order(-Count),]

Expense_data <- Expense_data[1:3,]
print(Expense_data)
```
Three highest expenses are food, water, and electricity


**How many households have at least one member that requires medication/medical care on a regular basis? **
```{r}
#fixing minor error seen in the data
data$Q16[125] <- "A"
data$Q16_Optional[125] <- "ARVS"

#Convert letter to option
data$Q16[data$Q16 == "A"] <- "Yes"
data$Q16[data$Q16 == "B"] <- "No"

#code to print bar graph
ggplot(data, aes(x = Q16)) + geom_bar(fill="steelblue") + labs(x = "Does anyone in the home take medication?", y = "Count")

medication_table <- count(data, vars = Q16)
medication_table <-medication_table %>%
  rename(Medication = vars, Count = n) %>%
  mutate(Percent = (Count/total)*100)
print(medication_table)
```


**What is the most common medication used? **
```{r}
#unique(data$Q16_Optional)
#correct data
data$Q16_Optional[data$Q16_Optional == "NA"] <- NA


data <- data %>%
  mutate(Natural.Herbs = case_when(grepl('LOCAL HERBS',Q16_Optional) ~1)) %>%
  mutate(Blood.Pressure = case_when(grepl('HIGH BLOOD PRESSURE MEDICATION|PRESSURE',Q16_Optional) ~1)) %>%
  mutate(HIV = case_when(grepl('ART|HIV|ARUS|NTIRETROUIRAL MEDICATION|ARVS|ANTIRETROVIRAL MEDICATION|ARVIS|ANTIRETROVIRAL MEDICINE|ATIRCTROVIRAL MEDICATION',Q16_Optional) ~1)) %>%
  mutate(Diabetes = case_when(grepl('DIABETES|DIABATES MEDICINE',Q16_Optional) ~1)) %>%
  mutate(Asthma = case_when(grepl('ASTHMA MEDICATION|ATHEMA',Q16_Optional) ~ 1)) %>%
  mutate(Ulcer = case_when(grepl('ULCER', Q16_Optional) ~ 1)) %>%
  mutate(Unkown = case_when(grepl('NOT KNOWN', 'KNOW', 'MENTION', Q16_Optional) ~ 1)) %>%
  mutate(Herpes = case_when(grepl("HERPES", Q16_Optional) ~ 1)) %>%
  mutate(Heart.Condition = case_when(grepl("HEART", Q16_Optional)~ 1)) %>%
  mutate(Mental.Health = case_when(grepl("MENTAL", Q16_Optional) ~ 1)) %>%
  mutate(Eye.Condition = case_when(grepl("EYE", Q16_Optional) ~ 1)) %>%
  mutate(Cancer = case_when(grepl("CANCER", Q16_Optional) ~ 1)) %>%
  mutate(Pain = case_when(grepl("PAIN", Q16_Optional) ~ 1)) %>%
  mutate(Anemic = case_when(grepl("PHOLIC|FOLIC|SICKLE CELL", Q16_Optional) ~ 1)) %>%
  mutate(Malaria = case_when(grepl("MALARIA", Q16_Optional) ~ 1)) %>%
  mutate(Allergies = case_when(grepl("PREDILISOLIN", Q16_Optional) ~ 1)) 

  
Q16.Optional.data <- data %>%
  select(Natural.Herbs, Blood.Pressure,HIV,Diabetes,Asthma,Ulcer,Unkown,Herpes,Heart.Condition,Mental.Health,Eye.Condition,Cancer,Pain,Anemic,Malaria,Allergies)

Q16.Optional.data <- colSums(Q16.Optional.data, na.rm = TRUE)
print(Q16.Optional.data)
```


**What is the frequency of different illnesses in the sample population? **
```{r}
Q17_Options <-c("Q17.A", "Q17.B","Q17.C", "Q17.D", "Q17.E", "Q17.F","Q17.G")
for(x in Q17_Options){
  data[x] <- 0
}

for(i in 1:nrow(data)){
  sen <- strsplit(data$Q15[i], ",")
  data[i,"Q17.A"] <- ifelse(grepl("A", sen),1,0)
  data[i,"Q17.B"] <- ifelse(grepl("B", sen),1,0)
  data[i,"Q17.C"] <- ifelse(grepl("C", sen),1,0)
  data[i,"Q17.D"] <- ifelse(grepl("D", sen),1,0)
  data[i,"Q17.E"] <- ifelse(grepl("E", sen),1,0)
  data[i,"Q17.F"] <- ifelse(grepl("F", sen),1,0)
  data[i,"Q17.G"] <- ifelse(grepl("G", sen),1,0)
}

ill <- c("HIV", "Diabetes","Heart Condition", "Tuberculosis", "Malaria", "Cancer","Other")
count <- c(sum(data$Q17.A), sum(data$Q17.B), sum(data$Q17.C),sum(data$Q17.D),sum(data$Q17.E),sum(data$Q17.F),sum(data$Q17.G))

Q17_count <- do.call(rbind, Map(data.frame, Illness=ill, Frequency=count))

#add percentage of homes with illness
Q17_count <- Q17_count %>%
  mutate(Percentage = (Frequency/nrow(data))*100)

print(Q17_count)


ggplot(Q17_count, aes(x = Illness, y = Frequency)) + geom_bar(stat="identity", fill="steelblue") + labs(x = "Illness", y = "Count")+ ggtitle("Frequency of Illness Based on Raw Numbers")

ggplot(Q17_count, aes(x = Illness, y = Percentage)) + geom_bar(stat="identity", fill="steelblue") + labs(x = "Illness", y = "Percentage (%)")+ ggtitle("Frequency of Illness Based on Proportion")

```


**Does the education level of the respondent affect their knowledge of additional symptoms of COVID19? **
Note: Couldn't conduct any statistical analysis since it's very very difficult to code out and might not make sense
```{r}
# educational makeup of each option: A, B, C , D
Question18.Data <- data[,c("Q3", "Q18.A", "Q18.B", "Q18.C", "Q18.D")]

#table 
aggregate(. ~ Q3, Question18.Data, sum)


#graphs 
Question18.Data <- Question18.Data %>%
  mutate_at(vars(`Q18.A`:`Q18.D`), ~ifelse(. == 1, "YES", "NO"))

#Option A 
ggplot(data=Question18.Data, aes(x=Q18.A, fill = Q3)) + geom_bar(position="stack") + labs(x = "Did they choose option A?", y = "Count") + ggtitle("Breakdown of Question 18 Option A: Fever")


#Option B
ggplot(data=Question18.Data, aes(x=Q18.B, fill = Q3)) + geom_bar(position="stack") + labs(x = "Did they choose option B?", y = "Count") + ggtitle("Breakdown of Question 18 Option G: Dry Cough")


#Option C
ggplot(data=Question18.Data, aes(x=Q18.C, fill = Q3)) + geom_bar(position="stack") + labs(x = "Did they choose option C?", y = "Count") + ggtitle("Breakdown of Question 18 Option C: Shortness of Breath")

#Option D
ggplot(data=Question18.Data, aes(x=Q18.D, fill = Q3)) + geom_bar(position="stack") + labs(x = "Did they choose option D?", y = "Count") + ggtitle("Breakdown of Question 18 Option D: Dry Cough")
```


**What does the sample population think are the common methods in which COVID19 spreads?**
```{r}
data %>% 
  select(Q20.A:Q20.I) %>% 
  colSums(na.rm=TRUE) * 100 / nrow(data) # percent for each option
```


**What percent of the sample population knows the common methods in which COVID19 spreads?**
```{r}
data %>% 
  filter(Q20.B==1 & Q20.D==1 & Q20.E==1) %>% # sample that selected all of the correct answers ( B, D, E)
  nrow() * 100 / nrow(data) # compute their percent
```


**Does the sample population know important facts about COVID19? **
```{r}
false_cols <- c("Q21.1", "Q21.2", "Q21.4", "Q21.5", "Q21.6", "Q21.10")
data <- data  %>% 
  mutate(across(all_of(false_cols), function(x) {
    ifelse(x=="FALSE", 1, 0)
  }))

true_cols <- c("Q21.3", "Q21.7", "Q21.8", "Q21.9", "Q21.11")
data <- data  %>% 
  mutate(across(all_of(true_cols), function(x) {
    ifelse(x=="TRUE", 1, 0)
  }))

data <- data %>%
  mutate(across(Q22, function(x) {
  ifelse(x=="B", 1, 0)
  }))

data %>% 
  select(Q21.1:Q22, -Q21.4_Notes)  %>% 
  colSums(na.rm=TRUE) * 100 / nrow(data) # percent correct for each option
```


**Does the level of education of the respondent affect the COVID19 knowledge score?**
```{r}

data[, "knowledge_score"] <- data %>% 
  select(Q21.1:Q22, -Q21.4_Notes)  %>% 
  rowSums(na.rm=TRUE)

data %>%
  select(Q3, knowledge_score) %>%
  group_by(Q3) %>%
  summarize(count=n(),
            mean_score=mean(knowledge_score, na.rm = TRUE))

cor(data %>% 
         select(Q3, knowledge_score) %>% 
         mutate_if(is.factor, as.numeric))

library(GGally)
ggcorr(data %>% 
         select(Q3, Q21.1:Q22, -Q21.4_Notes, knowledge_score) %>% 
         mutate_if(is.factor, as.numeric), 
       label=TRUE, 
       label_size=2, 
       label_round=2, 
       label_alpha=TRUE)
```


**Does the sector in which in the household work in affect the respondent’s COVID19 knowledge score?**
```{r}
data %>% count(, sort=TRUE) # too many sectors, need to reduce this manually

data %>% 
  select(Q11_Source1, knowledge_score) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate_if(is.factor, as.numeric)

data %>% 
  select(agriculture:other)  %>% 
  colSums(na.rm=TRUE) * 100 / nrow(data)
```


**How many people think that handwashing can protect against COVID19? **
```{r}
data %>%
  select(Q22) %>%
  group_by(Q22) %>%
  summarize(percent_score=n() * 100 / nrow(data))
```


**Does the sample population believe that COVID19 could be treated during the survey period? **
```{r}
data %>% count(Q23.1) # need to coerce values
```

**Does the sector in which in the household work in affect the respondent’s COVID19 knowledge score?**
```{r}

sector.lm <- lm(knowledge_score  ~ agriculture + transportation + education + sponsership + construction + business + trade + government + law + health + food + other, data = data)

kable(coef(summary(sector.lm)), digits = c(3,3,3,3))

```
CONCLUSION: Those in transportation seem to have a lower COVID-knowledge score at a 5% level 


**How many people think that handwashing can protect against COVID19? **
```{r}
data %>%
  select(Q22) %>%
  group_by(Q22) %>%
  summarize(percent_score=n() * 100 / nrow(data))
```

**How does handwashing protect against COVID19 (if they selected “yes”)**

```{r}
Q23_table <- Q22_table[-c(4,5),] #remove other options
print(Q23_table) #printed chart since numbers are small

ggplot(Q23_table, aes(x = Option, y = Percentage)) + geom_bar(stat="identity", fill="steelblue") + labs(x = "Option", y = "Count")
```

**Does the sample population believe that COVID19 could be treated during the survey period? **

```{r}
Q23_1_table <- data %>%
  count(Q23.1)

Q23_1_table$Percentage <- (Q23_1_table$n/total)*100
Q23_1_table <- Q23_1_table %>%
  rename(Option = Q23.1, Count = n)

print(Q23_1_table)
```


**Why does the sample population think COVID19 can be treated/not treated? **
```{r}
#Broke up Qustion 17 based on the following options 

Q23_2_Options <-c("Q23.2.A", "Q23.2.B","Q23.2.C", "Q23.2.D", "Q23.2.E")
for(x in Q23_2_Options){
  data[x] <- 0
}

for(i in 1:nrow(data)){
  sen <- strsplit(data$Q15[i], ",")
  data[i,"Q23.2.A"] <- ifelse(grepl("A", sen),1,0)
  data[i,"Q23.2.B"] <- ifelse(grepl("B", sen),1,0)
  data[i,"Q23.2.C"] <- ifelse(grepl("C", sen),1,0)
  data[i,"Q23.2.D"] <- ifelse(grepl("D", sen),1,0)
  data[i,"Q23.2.E"] <- ifelse(grepl("E", sen),1,0)
}

Q23_2_Data <- data %>%
  select(Q23.2.A, Q23.2.B,Q23.2.C, Q23.2.D, Q23.2.E)

Q23_2_Data <- melt(Q23_2_Data, measure.vars = c("Q23.2.A", "Q23.2.B","Q23.2.C", "Q23.2.D", "Q23.2.E"))

Q23_2_Data <- Q23_2_Data %>% 
  rename(Option = variable, Value = value) %>%
  group_by(Option) %>%
  summarise(Total = sum(Value))

ggplot(data=Q23_2_Data, aes(x=Option, y=Total)) + geom_bar(stat="identity", fill = "royalblue4") + labs(x = "Option", y = "# of households") + ggtitle("# of households who chose this option")

print(Q23_2_Data)
```
**What is the respondent’s education level for each selection?**
```{r}
Q23__Data <- data %>%
  select(Q3, Q23.1, Q23.2.A, Q23.2.B,Q23.2.C, Q23.2.D, Q23.2.E)

Q23_1_BarChart <- Q23__Data %>%
  select(Q3, Q23.1)

#Plot for Question 23.1
ggplot(data=Q23_1_BarChart, aes(x=Q23.1, fill = Q3)) + geom_bar(position="stack") + labs(x = "Choice", y = "Count") + ggtitle("Breakdown of Question 23.1 by Education")

#table for Option 1
print(Q23_1_BarChart)

Q23_2_BarChart <- Q23__Data %>%
  select(Q3, Q23.2.A, Q23.2.B,Q23.2.C, Q23.2.D, Q23.2.E)

Q23_2_BarChart <- melt(Q23_2_BarChart, measure.vars = c("Q23.2.A", "Q23.2.B","Q23.2.C", "Q23.2.D", "Q23.2.E"))

Q23_2_BarChart <- Q23_2_BarChart %>%
  group_by(Q3, variable)%>%
  summarise(value = sum(value))

#Plot for Question 23.2
ggplot(data=Q23_2_BarChart, aes(x=variable,y = value, fill = Q3)) + geom_bar(position="stack", stat="identity") + labs(x = "Choice", y = "Count") + ggtitle("Breakdown of Question 23.2 by Education")

#table breakdown
print(Q23_2_BarChart)
```

**What household economic statuses make up each selection?**
```{r}
#quantile(data$Q13, na.rm = TRUE) <- used to check the quantile breakdown

#Create quantile category for each dataset based on output for quantile
data$Quantile <- 0
data$Quantile <- ifelse(data$Q13 <= 9000, "Low", 
              ifelse(data$Q13 > 9000 & data$Q13 <= 20000, "Medium", "High"))
data$Quantile <- factor(data$Quantile, levels = c("Low", "Medium", "High"))

Economic_Q23_Data <- data %>%
  select(Quantile,Q23.1, Q23.2.A, Q23.2.B,Q23.2.C, Q23.2.D, Q23.2.E )

#Plot for Question 23.1
ggplot(data=Economic_Q23_Data, aes(x=Q23.1, fill = Quantile)) + geom_bar(position="stack") + labs(x = "Choice", y = "Count") + ggtitle("Breakdown of Question 23.1 by Economic Range")

#table for Question 1
print(Economic_Q23_Data)

Economic_Q23_2_Data <- Economic_Q23_Data %>%
  select(Quantile, Q23.2.A, Q23.2.B,Q23.2.C, Q23.2.D, Q23.2.E)

Economic_Q23_2_Data <- melt(Economic_Q23_2_Data, measure.vars = c("Q23.2.A", "Q23.2.B","Q23.2.C", "Q23.2.D", "Q23.2.E"))

Economic_Q23_2_Data <- Economic_Q23_2_Data %>%
  group_by(Quantile, variable)%>%
  summarise(value = sum(value))

#Plot for Question 23.2
ggplot(data=Economic_Q23_2_Data, aes(x=variable,y = value, fill = Quantile)) + geom_bar(position="stack", stat="identity") + labs(x = "Choice", y = "Count") + ggtitle("Breakdown of Question 23.2 by Education")

#table breakdown
print(Economic_Q23_2_Data)

```

Note: 
Quantile Breakdown is the following: 
0%: 3000
25%: 9000
50%: 14000
75%: 20000
100%: 340000

**What are the top 3 household supplies used to protect against COVID19? **
Note: Some mentioned social distancing but that is not a supply so I did not include it here. 

Note: Jerryau is the yellow jug people can carry water and other liquids in
```{r}
#function to take list into strings
data <- data %>%
  mutate(Masks = case_when(Q24 %like% 'MASK|FACE MASK|MASKS' ~1)) %>%
  mutate(Soap = case_when(Q24 %like% 'SOAP|SOAPS' ~1)) %>%
  mutate(Water = case_when(Q24 %like% 'WATER' ~1)) %>%
  mutate(Bathroom.Appliance = case_when(Q24 %like% 'FACILITY|BASIN' ~1)) %>%
  mutate(Sanitizer = case_when(Q24 %like% 'SANITIZER' ~1)) %>%
  mutate(Vegetables = case_when(Q24 %like% 'VEGETABLES|GREENS' ~1)) %>%
  mutate(Herbs = case_when(Q24 %like% 'HERBS|HERBAL' ~1)) %>%
  mutate(Fruits = case_when(Q24 %like% 'FRUITS|LEMON' ~1)) %>%
  mutate(alcohol = case_when(Q24 %like% 'ALCOHOL' ~1)) %>%
  mutate(Jerryau = case_when(Q24 %like% 'JERRYAU|JERRYCAN|JERRYCAU' ~1)) %>%
  mutate(Food = case_when(Q24 %like% 'SOUP|GHARRIC|SALAMI' ~1)) 


Supplies <- data %>%
  select(Masks, Soap, Water, Bathroom.Appliance, Sanitizer, Vegetables, Herbs, Fruits, alcohol, Jerryau, Food)

Supplies <- melt(Supplies, measure.vars = c("Masks","Soap", "Water", "Bathroom.Appliance", "Sanitizer", "Vegetables", "Herbs", "Fruits", "alcohol", "Jerryau", "Food"))

Supplies <- Supplies %>%
  group_by(variable) %>%
  mutate(Value = sum(value, na.rm = TRUE))

Supplies <- distinct(select(Supplies, -value), variable, Value)

Supplies <- Supplies %>%
  mutate(Percentage = (Value/nrow(data))*100) %>%
  rename(Purchases = variable)

print(Supplies)


```

**What are the top 3 precautions taken to protect against COVID19?**
```{r}

data <- data %>%
  mutate(Wearing.Masks = case_when(grepl('MASK|FACE MASK',Q25) ~1)) %>%
  mutate(Washing = case_when(grepl('HANDS|WASHING|WASH',Q25) ~1)) %>%
  mutate(Stay.at.home = case_when(grepl('HOME',Q25) ~1)) %>%
  mutate(Social.Distancing = case_when(grepl('SOCIAL|DISTANCE|OCIAL',Q25) ~1)) %>%
  mutate(Not.Touching = case_when(grepl('HUGS|HANDSHAKE|HUG|TOUCH|HAND SHAKE',Q25) ~1)) %>%
  mutate(Sanitze = case_when(grepl('SANITIZING|SANITIZE|SANITIZER',Q25) ~1)) %>%
  mutate(Herbs = case_when(grepl('HERBAL|HERB|HERBS',Q25) ~1)) %>%
  mutate(Vegetables = case_when(grepl('VEGETABLES',Q25) ~1)) %>%
  mutate(Fruits = case_when(grepl('FRUITS',Q25) ~1)) %>%
  mutate(Stay.in.Garden = case_when(grepl('GARDEN',Q25) ~1)) %>%
  mutate(Keep.Home.Clean = case_when(grepl('HOME & CLEAN',Q25) ~1)) %>%
  mutate(wear.gloves = case_when(grepl('GLOVE',Q25) ~1)) %>%
  mutate(not.allowing.people.into.the.house = case_when(grepl('VISITORS' ,Q25) ~1)) %>%
  mutate(not.allowing.people.into.the.house = case_when(grepl('COMMUNITIES|COMMUNITY|MEETINGS|GATHERING|PARTIES' ,Q25) ~1)) %>%
  mutate(drink.warm.water = case_when(grepl('WARM WATER' ,Q25) ~1))


Precautions <- data %>%
  select(Wearing.Masks, Washing, Stay.at.home, Social.Distancing, Not.Touching, Sanitze, Herbs, Fruits, Vegetables, Stay.in.Garden, Keep.Home.Clean, wear.gloves, not.allowing.people.into.the.house)

Precautions <- melt(Precautions, measure.vars = c("Wearing.Masks","Washing", "Stay.at.home", "Social.Distancing", "Not.Touching", "Sanitze", "Herbs", "Fruits", "Vegetables", "Stay.in.Garden", "Keep.Home.Clean", "wear.gloves", "not.allowing.people.into.the.house"))

Precautions <- Precautions %>%
  group_by(variable) %>%
  mutate(Value = sum(value, na.rm = TRUE))

Precautions <- distinct(select(Precautions, -value), variable, Value)

Precautions <- Precautions %>%
  mutate(Percentage = (Value/nrow(data))*100) %>%
  rename(Purchases = variable)

print(Precautions)

```

**What are the top 3 actions taken in the event that someone in the household shows symptoms of COVID19?**
Note: Some people in the community seem to have been given toll free numbers to call medical workers in the event someone has COVID
Note: Some used words vageuly such as "leaders" or "authorites." Not sure who they are for sure so they were placed in "leaders" 
Note: One person said they don't know what the symptoms of COVID-19 are

```{r}
data <- data %>%
  mutate(Avoid.person = case_when(grepl('RUN AWAY|DISTANCE', Q26) ~ 1)) %>%
  mutate(Contact.Ministry.of.Health = case_when(grepl('MINISTRY', Q26) ~ 1)) %>%
  mutate(Go.to.hospital = case_when(grepl('HOSPITAL|FACILITY', Q26) ~ 1)) %>%
  mutate(inform.LCI = case_when(grepl('L.C.I|LCI', Q26)~ 1)) %>%
  mutate(Inform.Health.Worker = case_when(grepl("HEALTH|MEDIC", Q26) ~ 1)) %>%
  mutate(keep.at.home = case_when(grepl("HOME", Q26) ~ 1)) %>%
  mutate(call.police = case_when(grepl("POLICE", Q26) ~ 1)) %>%
  mutate(call.VHT = case_when(grepl("VHT", Q26) ~ 1)) %>%
  mutate(Dont.know = case_when(grepl("KNOW|SURE", Q26) ~ 1)) %>%
  mutate(Isolating.infected.person = case_when(grepl("ISOLAT", Q26) ~1)) %>%
  mutate(give.herbal.medicine = case_when(grepl("HERB", Q26) ~ 1)) %>%
  mutate(inform.local.leaders = case_when(grepl("LEADERS|AUTHORITY|GOVERNMENT", Q26) ~ 1)) %>%
  mutate(juice = case_when(grepl("JUICE", Q26) ~ 1)) %>%
  mutate(COVID.task.force = case_when(grepl("TASK", Q26) ~ 1)) %>%
  mutate(Inform.LCI.Chairperson = case_when(grepl("CHAIR", Q26) ~ 1))
  
Actions <- data %>%
  select(Avoid.person, Contact.Ministry.of.Health, Go.to.hospital, inform.LCI, Not.Touching, Inform.Health.Worker, keep.at.home, call.police, call.VHT, Dont.know, Isolating.infected.person, give.herbal.medicine,inform.local.leaders, juice, COVID.task.force, Inform.LCI.Chairperson)

Actions <- melt(Actions, measure.vars = c("Avoid.person","Contact.Ministry.of.Health", "Go.to.hospital", "inform.LCI", "Not.Touching", "Inform.Health.Worker", "keep.at.home", "call.police", "call.VHT", "Dont.know", "Isolating.infected.person", "give.herbal.medicine", "inform.local.leaders", "juice", "COVID.task.force", "Inform.LCI.Chairperson"))

Actions <- Actions %>%
  group_by(variable) %>%
  mutate(Value = sum(value, na.rm = TRUE))

Actions <- distinct(select(Actions, -value), variable, Value)

Actions <- Actions %>%
  mutate(Percentage = (Value/nrow(data))*100) %>%
  rename(Purchases = variable)

  
print(Actions)
  
```

**What are the top 3 actions taken in the event that someone in the household was in a place with people that might have COVID19?**

```{r}
Q27.data <- data %>%
  mutate(Avoid.person.or.place = case_when(grepl('AWAY|DISTANCE|AVOID', Q27) ~ 1)) %>%
  mutate(Contact.Ministry.of.Health = case_when(grepl('MINISTRY', Q27) ~ 1)) %>%
  mutate(Go.to.hospital = case_when(grepl('HOSPITAL|FACILITY', Q27) ~ 1)) %>%
  mutate(inform.LCI = case_when(grepl('L.C.I|LCI', Q27)~ 1)) %>%
  mutate(Inform.Health.Worker = case_when(grepl("HEALTH|MEDIC", Q27) ~ 1)) %>%
  mutate(keep.at.home = case_when(grepl("HOME", Q27) ~ 1)) %>%
  mutate(call.police = case_when(grepl("POLICE", Q27) ~ 1)) %>%
  mutate(call.VHT = case_when(grepl("VHT", Q27) ~ 1)) %>%
  mutate(Dont.know = case_when(grepl("KNOW|SURE", Q27) ~ 1)) %>%
  mutate(Isolating.infected.person = case_when(grepl("ISOLAT|QUARAN", Q27) ~1)) %>%
  mutate(give.herbal.medicine = case_when(grepl("HERB", Q27) ~ 1)) %>%
  mutate(inform.local.leaders = case_when(grepl("LEADERS|AUTHORITY|GOVERNMENT", Q27) ~ 1)) %>%
  mutate(juice = case_when(grepl("JUICE", Q27) ~ 1)) %>%
  mutate(COVID.task.force = case_when(grepl("TASK", Q27) ~ 1)) %>%
  mutate(Inform.LCI.Chairperson = case_when(grepl("CHAIR", Q27) ~ 1)) %>%
  mutate(Wash = case_when(grepl("WASH", Q27) ~ 1)) %>%
  mutate(Get.tested = case_when(grepl("TEST", Q27) ~ 1)) %>%
  mutate(inform.RDC = case_when(grepl("RDC", Q27)~ 1))

Actions2 <- Q27.data %>%
  select(Avoid.person, Contact.Ministry.of.Health, Go.to.hospital, inform.LCI, Not.Touching, Inform.Health.Worker, keep.at.home, call.police, call.VHT, Dont.know, Isolating.infected.person, give.herbal.medicine,inform.local.leaders, juice, COVID.task.force, Inform.LCI.Chairperson, Wash, Get.tested, inform.RDC)

Actions2 <- melt(Actions2, measure.vars = c("Avoid.person","Contact.Ministry.of.Health", "Go.to.hospital", "inform.LCI", "Not.Touching", "Inform.Health.Worker", "keep.at.home", "call.police", "call.VHT", "Dont.know", "Isolating.infected.person", "give.herbal.medicine", "inform.local.leaders", "juice", "COVID.task.force", "Inform.LCI.Chairperson", "Wash", "Get.tested", "inform.RDC"))

Actions2 <- Actions2 %>%
  group_by(variable) %>%
  mutate(Value = sum(value, na.rm = TRUE))

Actions2 <- distinct(select(Actions2, -value), variable, Value)

Actions2 <- Actions2 %>%
  mutate(Percentage = (Value/nrow(data))*100) %>%
  rename(Purchases = variable)

  
print(Actions2)

```

**What is the relationship between the source of information and the COVID19 knowledge score?**

```{r}
Q28_data <- data %>%
  select(Q28.A, Q28.B, Q28.C, Q28.D, Q28.E, Q28.F, Q28_Other, knowledge_score)

Q28_data <- Q28_data %>%
  mutate(TV = case_when(Q28_Other %like% 'TV|TELE|TELEVISION|YOUTUBE' ~ 1)) %>%
  mutate(Government = case_when(Q28_Other %like% 'HEALTH|POLIT|MEDICAL|MEDICINE' ~ 1)) %>% mutate(Q28.C = case_when(Q28_Other %like% 'FRIENDS|FRI|PEOPLE|COMMUNITY' | Q28.C == 1~ 1)) %>% mutate(Q28.B = case_when(Q28_Other %like% 'NEWSPAPER|NEW|NEWS' | Q28.B == 1~ 1)) %>% mutate(Q28.E = case_when(Q28_Other %like% 'PHONE' | Q28.E == 1~ 1) ) %>% rename(Radio = Q28.A, Newsletter = Q28.B, Friends = Q28.C, Religious.figures = Q28.D, SocialMedia = Q28.E)

Q28_data <- select(Q28_data, -Q28_Other)
Q28_data[is.na(Q28_data)] <- 0

Source_Knowledge <- aov(knowledge_score ~ Radio + Newsletter + Friends + Religious.figures + SocialMedia + TV + Government, data = Q28_data)

summary(Source_Knowledge)
  

```

**What are the sources of COVID19 information for the sample population?**

```{r}

Q28_data <- select(Q28_data, -c(Q28.F, knowledge_score))

Q28_melt <- melt(Q28_data, measure.vars = c("Radio", "Newsletter", "Friends", "Religious.figures", "SocialMedia", "TV", "Government"))

Q28_melt <- Q28_melt %>%
  group_by(variable) %>%
  mutate(Value = sum(value))

Q28_melt <- distinct(select(Q28_melt, -value), variable, Value)

Q28_melt <- Q28_melt %>%
  mutate(Percentage = (Value/nrow(data))*100)

ggplot(data=Q28_melt, aes(x=variable, y=Value)) + geom_bar(stat="identity", fill = "royalblue4") + labs(x = "Source of Information", y = "# of households") + ggtitle("# of households who use each source for information")

ggplot(data=Q28_melt, aes(x=variable, y=Percentage)) + geom_bar(stat="identity", fill = "royalblue4") + labs(x = "Source of Information", y = "Percentage of households") + ggtitle("% of households who use each source for information")

print(Q28_melt)

```

**Is there a correlation between the number of times COVID19 news is obtained per week and education?**
```{r}
COVID_education.lm <- glm(Q29 ~ Q3, data = data)

kable(coef(summary(COVID_education.lm)), digits = c(3,3,3,3))
```


**Is there a correlation between the number of times COVID19 news is obtained per week and economic status?**

```{r}
COVID_economic.lm <- glm(Q29 ~ Quantile, data = data)

kable(coef(summary(COVID_economic.lm)), digits = c(3,3,3,3))

```
**Is there a correlation between the number of times  COVID19 news is obtained per week and age of respondent?**

```{r}
COVID_age.lm <- glm(Q29 ~ Age_of_Respondent, data = data)
kable(coef(summary(COVID_age.lm)), digits = c(3,3,3,3))
```
**What is the (average/median) number of times  COVID19 news is obtained per week?**

```{r}
Q29_vector <- data$Q29
#calculate mean
Q29_mean <- mean(Q29_vector, na.rm = TRUE) #remove anyone who didn't answer this question
#calculate median
Q29_med <- median(Q29_vector, na.rm = TRUE) #remove anyone who didn't answer this question
#calculate mode
Q29_mode <- mode(Q29_vector)

#calculate range
Q29_range <- range(Q29_vector, na.rm = TRUE) #remove anyone who didn't answer this

print(Q29_mean) #mean
print(Q29_med) #median
print(Q29_mode) #mode
quantile(Q29_vector, na.rm = TRUE) #quartiles
```

**What are the measures taken by households to limit the spread of COVID19?**
```{r}
Q30_data <- data %>%
  select(Q30.A, Q30.B, Q30.C, Q30.D, Q30.E)
  
Q30_melt <- melt(Q30_data, measure.vars = c("Q30.A","Q30.B", "Q30.C", "Q30.D", "Q30.E"))

Q30_melt <- Q30_melt %>%
  group_by(variable) %>%
  mutate(Value = sum(value))

Q30_melt <- distinct(select(Q30_melt, -value), variable, Value)

Q30_melt <- Q30_melt %>%
  mutate(Percentage = (Value/nrow(data))*100) %>%
  rename(Variable = variable)


ggplot(data=Q30_melt, aes(x=Variable, y=Percentage)) + geom_bar(stat="identity", fill = "royalblue4") + labs(x = "Source of Information", y = "Percentage of households") + ggtitle("% of households who use each source for information")

print(Q30_melt)
  
```

**How do households handle children interacting with people from other households? Are they handling them correctly?**
Note: Some families are recieving "COVID-19 Packges" though it's not clear what that is
Note: Ginny instructed to breakdown those who are doing the right thing and others are not. For the most part, most know what to do but the problem is that many parents mentioned that the kids don't want to and A. the kids are too old to be controlled, B. the parents are sick, C. They live in slums so socially distancing is not possible or D. there are no COVID cases in the area so they let the kids out. 


```{r}
Q31_data <- data %>%
select(Q31_STAY_AT_HOME, Q31_Not_Possible_to_keep_indoors,Q31_COVID_PACKAGES,Q31_Sanitize, Q31_Keep_Away_From_Others, Q31_Keep_Busy,Q31_Wash_Hands, Q31_Talks_to_kids_about_why_to_socially_distance,Q31_Garden,Q31_Cant_make_kids_stay_at_home, Q31_live_in_slums,Q31_Nothing,Q31_Too_sick_to_control_kids,Q31_Not_Sure,`Q31_God/Prayers`)

Q31_melt <- melt(Q31_data, measure.vars = c("Q31_STAY_AT_HOME","Q31_Not_Possible_to_keep_indoors", "Q31_COVID_PACKAGES", "Q31_Sanitize", "Q31_Keep_Away_From_Others","Q31_Keep_Busy", "Q31_Wash_Hands","Q31_Talks_to_kids_about_why_to_socially_distance", "Q31_Garden","Q31_Cant_make_kids_stay_at_home", "Q31_live_in_slums","Q31_Nothing", "Q31_Too_sick_to_control_kids","Q31_Not_Sure", "Q31_God/Prayers"))

Q31_melt <- Q31_melt %>%
  group_by(variable) %>%
  mutate(Value = sum(value, na.rm= TRUE))

Q31_melt <- distinct(select(Q31_melt, -value), variable, Value)

Q31_melt <- Q31_melt %>%
  mutate(Percentage = (Value/nrow(data))*100) %>%
  rename(Variable = variable)

print(Q31_melt)


```

**Can social distancing rules be followed in common public locations?**

```{r}
#Home
Home_data <- data %>%
  select(Q32_Home) %>%
  group_by(Q32_Home) %>%
  tally() %>%
  rename(Count = n) %>%
  mutate(Percentage = (Count/nrow(data))*100)

ggplot(Home_data, aes(x = Q32_Home, y = Percentage)) + geom_bar(stat="identity", fill="steelblue") + labs(x = "Answer", y = "% of Households") + ggtitle("Percentage of households who can social distance at home")

print(Home_data)

#Workplace
Work_data <- data %>%
  select(Q32_Workplace) %>%
  group_by(Q32_Workplace) %>%
  tally() %>%
  rename(Count = n) %>%
  mutate(Percentage = (Count/nrow(data))*100)

ggplot(Work_data, aes(x = Q32_Workplace, y = Percentage)) + geom_bar(stat="identity", fill="steelblue") + labs(x = "Answer", y = "% of Households") + ggtitle("Percentage of households who can social distance at the workplace")

print(Work_data)

#Food Markets
FoodMarkets_data <- data %>%
  select(Q32_FoodMarkets) %>%
  group_by(Q32_FoodMarkets) %>%
  tally() %>%
  rename(Count = n) %>%
  mutate(Percentage = (Count/nrow(data))*100)

ggplot(FoodMarkets_data, aes(x = Q32_FoodMarkets, y = Percentage)) + geom_bar(stat="identity", fill="steelblue") + labs(x = "Answer", y = "% of Households") + ggtitle("Percentage of households who can social distance at food markets")

print(FoodMarkets_data)

#Health Clinics
HealthClinic_data <- data %>%
  select(Q32_HealthClinics) %>%
  group_by(Q32_HealthClinics) %>%
  tally() %>%
  rename(Count = n) %>%
  mutate(Percentage = (Count/nrow(data))*100)

ggplot(HealthClinic_data, aes(x = Q32_HealthClinics, y = Percentage)) + geom_bar(stat="identity", fill="steelblue") + labs(x = "Answer", y = "% of Households") + ggtitle("Percentage of households who can social distance at Health Clinics")

print(HealthClinic_data)
#Religious Gatherings
Gatherings_data <- data %>%
  select(Q32_ReligiousGatherings) %>%
  group_by(Q32_ReligiousGatherings) %>%
  tally() %>%
  rename(Count = n) %>%
  mutate(Percentage = (Count/nrow(data))*100)

ggplot(Gatherings_data, aes(x = Q32_ReligiousGatherings, y = Percentage)) + geom_bar(stat="identity", fill="steelblue") + labs(x = "Answer", y = "% of Households") + ggtitle("Percentage of households who can social distance at Religious Gatherings")

print(Gatherings_data)

```

**Do most respondents cover their faces when they go outside?**
```{r}
#quick data cleaning
data$Q33[data$Q33 == "A"] <- "YES"

Q33_data <- data %>%
  select(Q33)


Q33_data <- Q33_data %>%
  group_by(Q33) %>%
  tally() %>%
  mutate(Percentage = (n/nrow(data)*100))

ggplot(data=Q33_data, aes(x=Q33, y=Percentage)) + geom_bar(stat="identity", fill = "royalblue4") + labs(x = "Do respondents cover their faces when they go outside?", y = "Percentage of households") + ggtitle("% of households who cover their faces when they go outside")

print(Q33_data)

```

**What did they use to cover their face?**

```{r}
Q33_data2 <- data %>%
  select(Q33, Q33_YES)

Q33_data2 <- Q33_data2 %>%
  mutate(Masks = case_when(Q33_YES %like% 'MASK' ~ 1)) %>%
  mutate(Cloth = case_when(Q33_YES %like% 'CLOTH|CHEIF|VEIL' ~1))

Q33_data2 <- Q33_data2 %>%
  filter(Q33 == "YES") %>%
  select("Masks", "Cloth")

Q33_melt <- melt(Q33_data2, measure.vars = c("Masks", "Cloth"))
Q33_melt <- na.omit(Q33_melt)

Q33_melt <- Q33_melt %>%
  group_by(variable) %>%
  tally() %>%
  rename(Type = variable, Count = n) %>%
  mutate(Percentage = Count/nrow(Q33_data2))
  
ggplot(data=Q33_melt, aes(x=Type, y=Percentage)) + geom_bar(stat="identity", fill = "royalblue4") + labs(x = "Type of Mask", y = "Percentage of households") + ggtitle("% of households to use masks based on material")

print(Q33_data)

```
**How important does the sample population consider mental health relative to physical health?**

```{r}
data$Q34[data$Q34 == "A"] <- "More Important"
data$Q34[data$Q34 == "B"] <- "Somewhat Important"
data$Q34[data$Q34 == "C"] <- "Equally Important"
data$Q34[data$Q34 == "D"] <- "Somewhat Less Important"
data$Q34[data$Q34 == "E"] <- "Not Important at All"
data$Q34[data$Q34 == "F"] <- "Refused to Answer"

Q34_data <- data %>%
  select(Q34) %>%
  group_by(Q34) %>%
  tally() %>%
  mutate(Percentage = (n/nrow(data))*100)

ggplot(data=Q34_data, aes(x=Q34, y = Percentage)) + geom_bar(stat="identity", fill = "royalblue4") + labs(x = "Level") + ggtitle("How important is mental health compared to physical health?") + theme(axis.text.x = element_text(angle = 90))

print(Q34_data)

```

**How many households know someone who has tested positive for COVID19?**
```{r}
data$Q35[data$Q35 == "A"] <- "Yes"
data$Q35[data$Q35 == "B"] <- "No"
data$Q35[data$Q35 == "C"] <- "I don't know"

Q35_data <- data %>%
  select(Q35) %>%
  group_by(Q35) %>%
  tally() %>%
  mutate(Percentage = (n/nrow(data))*100)

ggplot(data=Q35_data, aes(x=Q35, y = Percentage)) + geom_bar(stat="identity", fill = "royalblue4") + labs(x = "Level") + ggtitle("How many people know someone with COVID?") 

print(Q35_data)

```

**Does knowing someone who has tested positive for COVID19 increase the COVID19 knowledge score? **

```{r}
positive.lm <- glm(knowledge_score ~ Q35, data = data)
kable(coef(summary(positive.lm)), digits = c(3,3,3,3))
```

## Sindu's personal analysis

Taking stuff from earlier but then addressing questions that may be in interest to us

**Does the district you live in impact your COVID knowledge score**

```{r}
Masaka_data <- subset(data, District == "MASAKA")
Kalungu_data <- subset(data, District == "KALUNGU")

t.test(Masaka_data$knowledge_score, Kalungu_data$knowledge_score)

#average Maska knowledge score
print(mean(Masaka_data$knowledge_score))

#average Kalungu knowledge score
print(mean(Kalungu_data$knowledge_score))




```

Assuming significance at the 5% level, district you live in does not impact the knowledge score. However, it is signitificant at the 10% level. Those living in Masaka have a higher score compared to Kalungu. Kalungu has about 27 households that have transportation as their occupation compared to Masaka's 8 households

**Does gender impact COVID knowledge score?** 
```{r}
Male_data <- subset(data, Female == "Male")
Female_data <- subset(data, Female == "Female")

t.test(Male_data$knowledge_score, Female_data$knowledge_score)
```

Assuming significance at the 5% level, gender does not impact the knowledge score

**What could impact COVID knowledge score**

```{r}
knowledge.glm <- lm(knowledge_score ~ District + Female + Age_of_Respondent + Q3 + Q6 + Quantile + Q16 + Q22 + Q23.1 + agriculture + transportation + education + sponsership + construction + business + trade + government + law + health + food + other + Q29, data = data )

output <- kable(coef(summary(knowledge.glm)), digits = c(3,3,3,3))
print(output)

```







**Is there a relationship between the depression score and the respondent’s education level, gender, household position, household size, and number of children (<18) in the household?**
```{r}
data <- data %>%
  mutate_at(vars(Q36.1:Q36.9, -contains('Notes')), ~.-1)

data[, 'depression_score'] <- data %>% 
  select(Q36.1:Q36.9, -contains('Notes')) %>% 
  rowSums(na.rm=TRUE)

data %>%
  select(Q3, depression_score) %>%
  group_by(Q3) %>%
  summarize(count=n(),
            mean_score=mean(depression_score, na.rm = TRUE))

data %>%
  select(Position_of_Respondent, depression_score) %>%
  group_by(Position_of_Respondent) %>%
  summarize(count=n(),
            mean_score=mean(depression_score, na.rm = TRUE))

cor(data %>% 
         select(Q3, Q6, depression_score) %>% 
         mutate_if(is.factor, as.numeric))

ggcorr(data %>% 
         select(Q3, Female, Q6, Q7, Q36.1:Q36.9, depression_score, -contains('Notes')) %>% 
         mutate_if(is.factor, as.numeric), 
       label=TRUE, 
       label_size=2, 
       label_round=2, 
       label_alpha=TRUE)
```


**Are there any additional comments with regards to the depression survey? **
```{r}
df <- data %>% 
  select(contains('Notes') & contains('Q36'))
  
df[rowSums(is.na(df)) != ncol(df), ]
```


**Is there a relationship between the depression score and the difficulty for completing daily activities?**
```{r}
data %>%
  select(Q37, depression_score) %>%
  group_by(Q37) %>%
  summarize(count=n(),
            mean_score=mean(depression_score, na.rm = TRUE))
```


**Does knowing someone who has tested positive for COVID19 increase anxiety?**
```{r}
data <- data %>%
  mutate_at(vars(Q38.1:Q38.7, -contains('Notes')), ~.-1)

data[, 'gad_score'] <- data %>%
  select(Q38.1:Q38.7, -contains('Notes'))  %>% 
  rowSums(na.rm=TRUE)

data %>%
  select(Q35, gad_score) %>%
  group_by(Q35) %>%
  summarize(count=n(),
            mean_score=mean(gad_score, na.rm = TRUE))
```


**Is there a relationship between the anxiety score and the respondent’s education level, gender, household position, household size, and number of children (<18) in the household?**
```{r}
cor(data %>% 
         select(Q3, Q6, gad_score) %>% 
         mutate_if(is.factor, as.numeric))

ggcorr(data %>% 
         select(Q3, Female, Q6, Q7, Q38.1:Q38.7, gad_score, -contains('Notes')) %>% 
         mutate_if(is.factor, as.numeric), 
       label=TRUE, 
       label_size=2, 
       label_round=2, 
       label_alpha=TRUE)
```


**Is there a relationship between the anxiety score and the difficulty for completing daily activities? **
```{r}
data %>%
  select(Q39, gad_score) %>%
  group_by(Q39) %>%
  summarize(count=n(),
            mean_score=mean(gad_score, na.rm = TRUE))
```


```{r}
data %>%
  select(Q40_MentalHealthCounselor, gad_score, depression_score)  %>%
  group_by(Q40_MentalHealthCounselor) %>%
  summarize(count=n(), 
            mean_depression_score=mean(depression_score, na.rm = TRUE),
            mean_anxiety_score=mean(gad_score, na.rm = TRUE))

data %>%
  select(Q40_SpiritualHealther, gad_score, depression_score)  %>%
  group_by(Q40_SpiritualHealther) %>%
  summarize(count=n(),
            mean_depression_score=mean(depression_score, na.rm = TRUE),
            mean_anxiety_score=mean(gad_score, na.rm = TRUE))

data %>%
  select(`Q40_Doctor/Nurse`, gad_score, depression_score)  %>%
  group_by(`Q40_Doctor/Nurse`) %>%
  summarize(count=n(),
            mean_depression_score=mean(depression_score, na.rm = TRUE),
            mean_anxiety_score=mean(gad_score, na.rm = TRUE))
```


```{r}
library(tm)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)

text <- data %>%
          select(Q41_Additional_Comments)
docs <- Corpus(VectorSource(text))
docs <- docs %>%
  tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(stripWhitespace)
docs <- tm_map(docs, content_transformer(tolower))
docs <- tm_map(docs, removeWords, stopwords("english"))
dtm <- TermDocumentMatrix(docs) 
matrix <- as.matrix(dtm) 
words <- sort(rowSums(matrix),decreasing=TRUE) 
df <- data.frame(word = names(words),freq=words)
set.seed(42)

wordcloud(words = df$word, freq = df$freq, min.freq = 3,           
          max.words=200, random.order=FALSE, rot.per=0.35,            
          colors=brewer.pal(8, "Dark2"))
```


**Mean Scores**
```{r}
data %>% 
  select(contains('score'), 'gad_score') %>%
  summary()
```




**Correlations on entire data**
```{r}

library(corrplot)
sig=0.5
# convert data to numeric in order to run correlations
# convert to factor first to keep the integrity of the data
# each value will become a number rather than turn into NA
df_cor <- data %>% mutate_if(is.character, as.factor)
df_cor <- df_cor %>% mutate_if(is.factor, as.numeric)

# run a correlation and drop the insignificant ones
corr <- cor(df_cor)
# prepare to drop duplicates and correlations of 1     
corr[lower.tri(corr, diag=TRUE)] <- NA 
# drop perfect correlations
# corr[corr == 1] <- NA 
# turn into a 3-column table
corr <- as.data.frame(as.table(corr))
# remove the NA values from above 
corr <- na.omit(corr) 
# select significant values  
corr <- subset(corr, abs(Freq) > sig) 
# sort by highest correlation
corr <- corr[order(-abs(corr$Freq)),] 
# print table
print(corr)
# turn corr back into matrix in order to plot with corrplot
mtx_corr <- reshape2::acast(corr, Var1~Var2, value.var="Freq")

# plot correlations visually
corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
```

